BERT Model Compression and Weight Sharing

### 2019-05-12 paper/paper.tex
Remove floating 'HEAD' in paper.tex

Deep Neurel Networks (DDNs) have evolved to the state-of-art for many tasks in
language modeling. In the meantime, they have become increasingly large in
size, due to both pre-trained embeddings and deeper networks. However, in
order to deploy the models on more resource-conscious devices efficiently,
such as mobile devices or even older personal laptops, as opposed to on cloud
virtual machines, it is crucial to reduce the size of the models.

@@

% Example Structure:
% \begin{itemize}
% \item What information does a non-expert need to know about the problem domain?
% \item What data exists for this problem?
% \item What are the challenges/opportunities inherent to the data? (High dimensional, sparse, missing data, noise, structure, discrete/continuous, etc?)
% \end{itemize}

@@

The benefits of training in this fashion have been empirically documented
% TODO: paper on KD doing well? 
\citep{hinton2015distilling}
and have theoretical motivation. Our goal of any network is to learn the
conditional distribution 
\[ p(y_i = k \mid \bm{x}_i).\]
In traditional training, we do this by observing the draws $y_i$ from this
distribution. However, to the extent that the teacher network prediction is a
good approximation of the true conditional distribution, the student network
can learn faster because it is being pushed toward predicting the de-noised
expectation.

@@

To provide a concrete example of how this might work, suppose BERT captures
from its pre-training that ``bad'', ``horrible,'' and ``terrible'' are
approximate synonyms. In a sentiment task, these three words appear in three
different sentences, two of which are negative, and one of which is positive.
A fine-tuned BERT may provide a lower likelihood of being positive to the
positive sentence than it would otherwise even if

% TODO: KD benefits

For our knowledge distillation, we make use of a fine-tuned BERT on a task as
our teacher network and train a simpler model to emulate the predictions use
\nameref{sec:model}

% TODO: talk about KD

@@

% TODO: talk about DeepTwist™

\begin{comment}
\begin{itemize}
\item How do you plan on training your parameters / inferring the
  states of your latent variables (MLE / MAP / Backprop / VI / EM / BP / ...)

\item What are the assumptions implicit in this technique? Is it an approximation or exact? If it is an approximation what bound does it optimize?

\item What is the explicit method / algorithm that you derive for learning these parameters?
\end{itemize}
\end{comment}

\begin{algorithm}
  \begin{algorithmic}
    \STATE{}
  \end{algorithmic}
  \caption{Your Pseudocode}
\end{algorithm}

@@

% \begin{itemize}
% \item What is the formal definition of your problem?
% \item What is the precise mathematical model you are using to represent it? In almost all cases this will use the probabilistic language from class, e.g.
%   \begin{equation}
%   z \sim {\cal N}(0, \sigma^2)\label{eq:1}
% \end{equation}
% But it may also be a neural network, or a non-probabilistic loss,
% \[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]

@@

% \item What are the parameters or latent variables of this model that you plan on estimating or inferring? Be explicit. How many are there? Which are you assuming are given? How do these relate to the original problem description?
% \end{itemize}

% \begin{figure}
%   \centering
%   \missingfigure[figheight=8cm]{}
%   \caption{\label{fig:diagram} This is a good place to include a diagram showing how your model works. Examples include a graphical model or a neural network block diagram.}
% \end{figure}

@@

\subsection{Knowledge Distillation}
For knowledge distillation experiments, we use various configurations of
the student network:
\paragraphdot{LSTM} We use a one-layer LSTM network with hidden dimension
$H=200$. We first embed the one-hot representation of tokens, and we feed
the embeddings through an LSTM. We then map the last hidden state of the
LSTM through a linear layer for classification. 
\paragraphdot{CNN} We use an implementation\footnote{\url{https://github.com/jfilter/cnn-text-classification-pytorch/blob/master/cnn_text_classification/model.py}} 
of \citet{kim2014convolutional} with kernel sizes $(3,4,5)$, hidden size
$H=300$, no dropout, and $20$ kernels for each kernel size.\footnote{These
are default in the implementation we used.}

@@

% \begin{table*}
%   \centering
%   \missingfigure{}
%   \caption{This is usually a table. Tables with numbers are generally easier to read than graphs, so prefer when possible.}
%   \label{fig:mainres}
% \end{table*}

\begin{table}[tbh]
        \caption{Results for DeepTwist}
        \label{tab:deeptwist_results}
        \centering
        \vspace{1em}
        \begin{tabular}{lccccc}
\toprule
{} &               type &  param &  val loss &   val acc \\
{} & 
\hypertarget{tabcol:1}{(1)} & \hypertarget{tabcol:2}{(2)} & \hypertarget{tabcol:3}{(3)} & \hypertarget{tabcol:4}{(4)} \\ \midrule
0  &  diff prune global &   $90$ &   $0.211$ &   $0.912$ \\
1  &              prune &   $20$ &   $0.291$ &   $0.886$ \\
2  &              prune &   $80$ &   $0.532$ &   $0.767$ \\
3  &         diff prune &   $80$ &   $0.222$ &   $0.916$ \\
4  &  diff prune global &   $99$ &       --- &       --- \\
5  &  diff prune global &   $90$ &   $0.213$ &   $0.916$ \\
6  &         diff prune &   $40$ &   $0.203$ &   $0.909$ \\
7  &                svd &   $50$ &   $0.490$ &   $0.810$ \\
8  &                svd &  $200$ &   $0.459$ &   $0.845$ \\
9  &              prune &   $40$ &   $0.424$ &   $0.845$ \\
10 &              prune &   $60$ &   $0.425$ &   $0.815$ \\
11 &  diff prune global &   $99$ &   $0.230$ &   $0.901$ \\
12 &  diff prune global &   $99$ &   $0.222$ &   $0.905$ \\
13 &         diff prune &   $99$ &   $0.222$ &   $0.913$ \\
14 &         diff prune &   $95$ &   $0.221$ &   $0.916$ \\
15 &         diff prune &   $90$ &   $0.216$ &   $0.916$ \\
16 &         diff prune &   $60$ &   $0.209$ &   $0.920$ \\
17 &                svd &  $100$ &   $0.515$ &   $0.823$ \\
18 &                svd &  $150$ &   $0.456$ &   $0.828$ \\
19 &         diff prune &   $80$ &   $2.533$ &   $0.491$ \\
20 &         diff prune &   $90$ &       --- &       --- \\
21 &  diff prune global &    $0$ &   $0.738$ &   $0.509$ \\
22 &         diff prune &    $0$ &   $0.692$ &   $0.522$ \\
23 &         diff prune &    $0$ &   $1.854$ &   $0.491$ \\
\bottomrule
\end{tabular}

@@

\section{Discussion}

>>>>

Deep Neural Networks (DDNs) have evolved to the state-of-art for many tasks in
language modeling. In the meantime, they have become increasingly large in
size, due to both pre-trained embeddings and deeper networks. However, in
order to deploy the models on more resource-conscious devices efficiently,
such as mobile devices or even older personal laptops, as opposed to on cloud
virtual machines, it is crucial to reduce the size of the models.

@@

% Example Structure:
% \begin{itemize}
% \item What information does a non-expert need to know about the problem domain?
% \item What data exists for this problem?
% \item What are the challenges/opportunities inherent to the data? (High
% dimensional, sparse, missing data, noise, structure, discrete/continuous,
% etc?)
% \end{itemize}

@@

The benefits of training in this fashion have been empirically documented
\citep{hinton2015distilling} and have theoretical motivation. Our goal of any
network is to learn the conditional distribution 
\[ p(y_i = k \mid \bm{x}_i).\]
In traditional training, we do this by observing the draws $y_i$ from this
distribution. However, to the extent that the teacher network prediction is a
good approximation of the true conditional distribution, the student network
can learn faster because it is being pushed toward predicting the de-noised
expectation.

@@

To provide a concrete example of how this might work, suppose BERT captures
from its pre-training that ``bad'', ``horrible,'' and ``terrible'' are
approximate synonyms. In a sentiment task, these three words appear in three
different sentences, two of which are negative, and one of which is positive.
A fine-tuned BERT may provide a lower likelihood of being positive to the
positive sentence than it would otherwise even if the word of interest appears
nowhere else.

For our knowledge distillation, we make use of a fine-tuned BERT on a task as
our teacher network and train a simpler model to emulate the predictions. For
the student network, we use both a CNN and an LSTM, specified further in
\cref{sec:model}.

@@

Knowledge distillation operates by passing knowledge from a teacher network to a
student network. There is no requirement that these networks resemble each other
in any way. We may, however, wish to directly compress a model whose
architecture we've had success with.

To this end, we follow the framework put forward by \citet{lee2018deeptwist},
\emph{DeepTwist}, whose training loop is characterized by \cref{algo:deeptwist}.
Training of the model closely resembles SGD with one important modification:
every $S_d$ batches, the model weights are transformed by a \emph{distortion}
function.

\begin{algorithm}
  \begin{algorithmic}
    \STATE{$n \leftarrow$ number of batches per epoch}
    \STATE{$\theta \leftarrow$ initialization of model parameters}
    \STATE{$k \leftarrow$ number of epochs to train}
    \STATE{$S_d \leftarrow$ distortion step}
    \FOR{$i=0$ to $k \cdot n$}
      \STATE{batch $\leftarrow$ batches[$i \pmod n$]}
      \STATE{$\theta \leftarrow \theta - \eta \nabla \ell(\theta,
      \text{batch})$}
      \COMMENT{gradient step}
      \IF{$i \equiv 0 \pmod{S_d}$}
        \STATE{$\theta \leftarrow \distortion(\theta)$}
      \ENDIF
    \ENDFOR
    \STATE{$\theta \leftarrow \distortion(\theta)$}
  \end{algorithmic}
  \caption{DeepTwist Training Loop}
  \label{algo:deeptwist}
\end{algorithm}

We have significant flexibility in picking a distortion function, but in order
to be useful, it must provide compression benefits. That is, the image of the
distortion function should have a parameterization that takes less space in
memory than then range of all possible $\theta$s. For example, if $\theta \in
\mathbb{R}^{m,n}$, and the distortion produces a matrix of rank 1, then any
distorted matrix could be uniquely identified by $\theta' \in \mathbb{R}^{n+m}$
through a matrix factorization.

It is important to note that while $\theta$ will be compressible immediately
after a distortion step, there is no guarantee that it will remain so after some
gradient steps. In fact, almost surely it will not be. Thus, we also enforce
that the process ends on a distortion step.

@@

% \begin{itemize}
% \item What is the formal definition of your problem?
% \item What is the precise mathematical model you are using to represent it? In
% almost all cases this will use the probabilistic language from class, e.g.
%   \begin{equation}
%   z \sim {\cal N}(0, \sigma^2)\label{eq:1}
% \end{equation}
% But it may also be a neural network, or a non-probabilistic loss,
% \[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]

@@

% \item What are the parameters or latent variables of this model that you plan
% on estimating or inferring? Be explicit. How many are there? Which are you
% assuming are given? How do these relate to the original problem description?
% \end{itemize}

% \begin{figure}
%   \centering
%   \missingfigure[figheight=8cm]{}
%   \caption{\label{fig:diagram} This is a good place to include a diagram
%   showing how your model works. Examples include a graphical model or a neural
%   network block diagram.}
% \end{figure}

@@

\subsection{Knowledge Distillation}

For knowledge distillation experiments, we use various configurations of the
student network: \paragraphdot{LSTM} We use a one-layer LSTM network with hidden
dimension $H=200$. We first embed the one-hot representation of tokens, and we
feed the embeddings through an LSTM. We then map the last hidden state of the
LSTM through a linear layer for classification.  \paragraphdot{CNN} We use an
implementation\footnote{\url{https://github.com/jfilter/cnn-text-classification-pytorch/blob/master/cnn_text_classification/model.py}}
of \citet{kim2014convolutional} with kernel sizes $(3,4,5)$, hidden size
$H=300$, no dropout, and $20$ kernels for each kernel size.\footnote{These are
default in the implementation we used.}

@@

% \begin{table*}
%   \centering
%   \missingfigure{}
%   \caption{This is usually a table. Tables with numbers are generally easier
%   to read than graphs, so prefer when possible.}
%   \label{fig:mainres}
% \end{table*}

% \begin{table}[tbh]
%         \caption{Results for DeepTwist}
%         \label{tab:deeptwist_results}
%         \centering
%         \vspace{1em}
%         \begin{tabular}{lccccc}
% \toprule
% {} &               type &  param &  val loss &   val acc \\
% {} & 
% \hypertarget{tabcol:1}{(1)} & \hypertarget{tabcol:2}{(2)} & \hypertarget{tabcol:3}{(3)} & \hypertarget{tabcol:4}{(4)} \\ \midrule
% 0  &  diff prune global &   $90$ &   $0.211$ &   $0.912$ \\
% 1  &              prune &   $20$ &   $0.291$ &   $0.886$ \\
% 2  &              prune &   $80$ &   $0.532$ &   $0.767$ \\
% 3  &         diff prune &   $80$ &   $0.222$ &   $0.916$ \\
% 4  &  diff prune global &   $99$ &       --- &       --- \\
% 5  &  diff prune global &   $90$ &   $0.213$ &   $0.916$ \\
% 6  &         diff prune &   $40$ &   $0.203$ &   $0.909$ \\
% 7  &                svd &   $50$ &   $0.490$ &   $0.810$ \\
% 8  &                svd &  $200$ &   $0.459$ &   $0.845$ \\
% 9  &              prune &   $40$ &   $0.424$ &   $0.845$ \\
% 10 &              prune &   $60$ &   $0.425$ &   $0.815$ \\
% 11 &  diff prune global &   $99$ &   $0.230$ &   $0.901$ \\
% 12 &  diff prune global &   $99$ &   $0.222$ &   $0.905$ \\
% 13 &         diff prune &   $99$ &   $0.222$ &   $0.913$ \\
% 14 &         diff prune &   $95$ &   $0.221$ &   $0.916$ \\
% 15 &         diff prune &   $90$ &   $0.216$ &   $0.916$ \\
% 16 &         diff prune &   $60$ &   $0.209$ &   $0.920$ \\
% 17 &                svd &  $100$ &   $0.515$ &   $0.823$ \\
% 18 &                svd &  $150$ &   $0.456$ &   $0.828$ \\
% 19 &         diff prune &   $80$ &   $2.533$ &   $0.491$ \\
% 20 &         diff prune &   $90$ &       --- &       --- \\
% 21 &  diff prune global &    $0$ &   $0.738$ &   $0.509$ \\
% 22 &         diff prune &    $0$ &   $0.692$ &   $0.522$ \\
% 23 &         diff prune &    $0$ &   $1.854$ &   $0.491$ \\
% \bottomrule
% \end{tabular}

% \end{table}

\input{../tabs/deeptwist_table.tex}

@@

\section{Discussion}
Things to say about deeptwist
\begin{itemize}
  \item Stratified prune better
  \item Prune rate mad robust
  \item Diff gud
  \item ``Improved performance''
  \item QQP takes long, but looks okay
  \item Didn't try only purning the $\bert$ part
\end{itemize}

Things to say about KD
\begin{itemize}
  \item it bad
  \item doens't work
  \item LSTM weird/CNN incrdibly noisy
  \item \TODO{compare}
\end{itemize}
